---
title: "Border West GP Outpatient Referral Appointments"
author: "Bradley Roy"
date: "2024-06-14"
output: html_document
---


## Background

The GP Cluster Profiles are sent to Cluster Quality Leads on a quarterly basis. In the quarter 2 2023 cluster report for <B>Borders West</B> there was a large increase in the number of GP outpatient referral appointments for Trauma. From quarter 2 2022 to quarter 2023, there was an increase of <B>+95.4%</B> (+158.9% for those aged 65 and over). This is a notably larger change compared to the "All Specialties" category which showed a decrease for <B>-7.2%</B> (-7.3% for those aged 65 and over). It was agreed that the increase in Trauma referrals warranted further consideration. 

This reports outlines GP Outpatient referral appointments for Borders West practices, over the period <B>1st March 2022 to 31st December 2023</B>. 


## Outpatient Analysis


### Borders West: Monthly GP Outpatient Referral Appointments by Age Group
```{r, echo=FALSE}
plotly_monthly_agegroup
```
Note: Specialties include "Dermatology", "General Surgery", "Trauma", "ENT"

### Borders West: Monthly GP Outpatient Referral Appointments by Speciality
```{r, echo=FALSE}
plotly_monthly_indicator
```

### Borders West: Monthly GP Trauma Outpatient Referral Appointments by Hospital
```{r, echo=FALSE}
plotly_monthly_hospital
```

### Borders West: GP Trauma Outpatient Referral Appointments for Quarter 2 by Practice
```{r, echo=FALSE}
plotly_total_practice
```

### Borders West: Monthly GP Trauma Outpatient Referral Appointments by Practice
```{r, echo=FALSE}
plotly_monthly_practice_numerator
```



